---
title: "Final project Draft1"
author: "Sakaiza"
date: "12/7/2020"
output: pdf_document
---

```{r, message= FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(bayesplot)
library(rstan)
library(rstanarm)
```


# Title

# Abstract


# Background and significance



# Methods

**Data collection**
```{r}
library(Lock5Data)
data(GPAGender)
```

**Variable creation**

**Analytic method**

## Framing the decision space




## The other information that helps us make a decision



## The loss function




## Uncertainity


# Results


# Discussion 


# References

=======
# Choosing high GPA students based on their SAT scores

# Abstract




# Background and significance



# Methods
In order to figure out the SAT score that maximizes a student's GPA below are the steps: 
Frame what the decision is
Set up fitted model predicting GPA based on Exercise, SAT and GenderCode
Establish a loss function that can maximize GPA values based on different SAT score values


**Data collection**

The data used in this project comes from the Lock5Data package in R. 
```{r}
library(Lock5Data)
data(GPAGender)

```

This data set contains 343 observations and 6 variables. 
Exercise: (hours per week)
SAT: out of 1600
GPA: 0-4 scale
Pulse: in beats per minute
Piercings: number of body piercings
GenderCode: 0 = female or 1 = male

**variables creation**

A variable called success was added, this codes 1 if the student's GPA is above or equal to  3.0 and 0 if it is below 3.0

```{r}
decision_data <- GPAGender%>%
  mutate(success = ifelse(GPA>=3.0,1,0))
view(decision_data)

```
## Framing the decision space

Here, we are trying to decide the SAT score that will determine successful admission of a student. The decision space, which is all the possible SAT scores, is then discrete. Let's denote the decision space as _S_ which means [0, 1600], and the decision _s_ $\epsilon$ [0, 1600]. 



## The other information that helps us make a decision
In Bayesian decision theory, an important is about choosing between all the different decision based on some information. 
Let's call that information $\theta$. This represents the probability that a student with a certain SAT score would be successful. $\theta$ $\epsilon$ [0,1]

## The loss function
This function will allow us to quantify how bad or good a given decision _s_ is given the information $\theta$. 
How good or bad the decision of choosing a student with a given SAT score, given the probability that they would succeed. 

L : $\theta$ x _s_ -> R

The loss function ties together the decision space and the information space. 
Since we want to:
maximize the SAT score at that which students will be successful.
minimize the SAT score at that which students will be successful.
In order to get a range. 


## Uncertainity

In order to express uncertainty in the deicsion, we will work with a probability distribution representing our belief in the true value of $\theta$, p($\theta$)
This is a distribution over $\theta$ conditionned on some observed data _x_. 

#code

```

# Results


# Discussion 


# References
