---
title: "Normal linear regression_draft"
author: "Chhiring Lama, Sakaiza Rasolofomanana Rajery"
date: "12/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(bayesplot)
library(rstan)
library(rstanarm)
```
# Explaining the association between exercising and academic performance



## Background

## Data set
```{r}
library(Lock5Data)
data(GPAGender)
```

##Analysis

**Variables**
In this model, we used the following variables:

Exercise (hours): Response variable, which is continuous
SAT (point) : Explanatory variable, which is continuous
GPA ( point): Explanatory variable, which is continuous

**Defining parameters**

Assuming that there is a linear association between exercise and SAT and GPA. The expected amount of exercise (Y) in hour for a student _i_ as a function of GPA (X) and SAT(Z) for that student ca be written as:

$Y_i$ = $\beta_0$ + $\beta_1$ * $X_i$ + $\beta_2$ *$Z_i$ + $\epsilon_i$ ~ N(0, $\sigma$)
where:

$\beta_0$ : The amount of exercise in hour when GPA and SAT score are both 0
$\beta_1$ : When SAT is kept constant, $\beta_1$ is the increase in hours of of exercise for every point increase in the student's GPA
$\beta_2$ : When GPA is kept constant, $\beta_2$ is the increase in hours of exercise for every point increase in the student's SAT
$\epsilon_i$: is a measure of the variability of Y for students with similar GPA and SAT scores. 

**Hypothesis**

We hypothesized that there is a negative association between hours of exercise and GPA, SAT. 

As GPA increases, the number of hours a student exercises decreases. Mathematically:
$H_0$ : $\beta_1$ = 0
$H_a$ : $\beta_1$ < 0

As SAT score increases, the number of hours a student exercise decreases. Mathematically:
$H_0$ : $\beta_2$ = 0
$H-a$ : $\beta_2$ < 0

**Building the Regression Model**

To identify the association between the number of hours exercised and academic progress (GPA and SAT score), a multiple linear regression was run with SAT score, GPA score and Pulse rate as the predictors and `Exercise` as the dependent variable. A bayesian linear regression model was created with **Rstan** package using the `stan_glm` function. The number of hours exercised (`Exercise`) could be predicted using `GPA` and `SAT`, assuming a normal (Gaussian) likelihood model. For this, model it is assumed that a normal-normal model can be used because, one can assume that the distribution of exercise as well as GPA and SAT scores is bell-shaped. The model contains an MCMC simulation with 4 Markov chains and with 10000 iterations for each chain. 




```{r}
set.seed(563453)

gpa_model_sim <- stan_glm(Exercise ~ SAT + GPA, data = GPAGender, family=  gaussian, chains = 4, iter = 10000*2)
``` 

**diagnosis**


```{R}
mcmc_trace(gpa_model_sim, size = 0.1)

```

```{r}
mcmc_dens_overlay(gpa_model_sim)

```

```{r}
gpa_model_sim <- summary(gpa_model_sim)

```

**Posterior credible interval**
```{R}
posterior_interval(gpa_model_sim, prob = 0.90)
mcmc_areas(gpa_model_sim,
           pars = c("(Intercept)"),
           prob = 0.90,
           prob = 0.90)
```


**posterior prediction**

```{r}
set.seed(563453)
r_prediction <- posterior_predict(
  gpa_model_sim,
  newdata = data.frame(GPA = 3.0)
)

mcmc_dens(r_prediction)
```

**One-  Test*