---
title: "Normal linear regression_draft"
author: "Chhiring Lama, Sakaiza Rasolofomanana Rajery"
date: "12/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning= FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(bayesplot)
library(rstan)
library(rstanarm)
library(reshape2)
library( janitor )
```


\centerline{\bfseries Explaining the association between exercising and academic performance}

## Background

When students focus on academics so much they tend to forget about their physical well being and exercise less. This analysis is an attempt to unveil the relationship between the exercising habit of students and their academic performance Using a data set of 343 variables from the Lock5Data Package, we will rely on bayesian methods, highlighting the importance of balancing prior information with data in order to infer a posterior interpretation,to explain the relationship between these two factors.   


## Data set
```{r}
library(Lock5Data)
data(GPAGender)

```

## Analysis


**Variables**


In this analysis, we used the following variables:

Exercise (hours): Response variable, which is continuous
SAT (point) : Explanatory variable, which is continuous
GPA ( point): Explanatory variable, which is continuous


**Defining the likelihood model and parameters**

Assuming that there is a linear association between exercise and SAT and GPA. The expected amount of exercise (Y) in hour for a student _i_ as a function of GPA (X) and SAT(Z) for that student can be written as:



\centerline {$Y_i = \beta_0 + \beta_1 * X_i + \beta_2 * Z_i + \epsilon_i ~ N(0, \sigma)$}

where:

\textbf {$\beta_0$} : The amount of exercise in hour when GPA and SAT score are both 0


\textbf {$\beta_1$} : The increase in hours of exercise for every point increase in the student's SAT, when GPA
is kept constant.


\textbf {$\beta_2$} : The increase in hours of of exercise for every point increase in the student's GPA, when SAT is kept constant


\textbf {$\epsilon_i$} : is a measure of the variability of Y for students with similar GPA and SAT scores. 


**Hypothesis**

We hypothesized that there is a negative association between hours of exercise and GPA, SAT. 

As SAT increases, the number of hours a student exercises decreases. Mathematically:

\textbf {$H_0$} : $\beta_1$ = 0

\textbf {$H_a$} : $\beta_1$ < 0 

As GPA score increases, the number of hours a student exercise decreases. Mathematically:

\textbf {$H_0$} : $\beta_2$ = 0 

\textbf {$H_a$} : $\beta_2$ < 0 

**Appropriate prior models**

Since $\beta_0$ can only take positive values and $\beta_1$ and $\beta_2$ can take real number values, a normal-normal model will be used to specify our prior understanding of their plausible values. 

Since the standard deviation $\sigma$ is always positive, an exponential model will be used to specify its prior. 

Our prior models will be as follows: 

$\beta_0$ ~ N(8,3^2)

$\beta_1$ ~ N(-0.0125, 0.0005^2) 

$\beta_2$ ~ N(-5,2^2) 

$\sigma$ ~ Exp(2) 


**Visualization**


```{r}
new_data <- GPAGender %>%
  gather(key, score, -c(Exercise, Pulse, Piercings, GenderCode))
  
ggplot(new_data, aes(score, Exercise))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  facet_wrap(~key,scales = "free_x" )

```
As shown in the figure above, in genereal, both GPA and SAT have negative linear correlation with exercise. That means that as the GPA and SAT scores increase, the predicted number of hours exercised per week is estimated to decrease. The decrease is steeper for GPA than SAT, so it can be assumed that the coefficient $\beta_1$ will be more negative than $\beta_2$. 

**Building the Regression Model**

To identify the association between the number of hours exercised and academic progress (GPA and SAT score), a multiple linear regression was run with SAT score, GPA score and Pulse rate as the predictors and `Exercise` as the dependent variable. 

A bayesian linear regression model was created with **Rstan** package using the `stan_glm` function. The number of hours exercised (`Exercise`) could be predicted using `GPA` and `SAT`, assuming a normal (Gaussian) likelihood model. For this, model it is assumed that a normal-normal model can be used because, one can assume that the distribution of exercise as well as GPA and SAT scores is bell-shaped. The model contains an MCMC simulation with 4 Markov chains and with 10000 iterations for each chain. 


```{r, message= FALSE, results='hide'}
set.seed(563453)

gpa_model_sim <- stan_glm(Exercise ~ SAT + GPA, data = GPAGender, family=  gaussian, chains = 4, iter = 10000*2)

``` 


**diagnosis** 

Once the model was created, diagnostics needed to be run to make sure that the markov chains used for the regression model was big enough (with adequate iterations) and provided more accurate approximation of the posterior. To check if the MCMC simulations meets the criteria, two plots were graphed: MCMC trace plot and density plot overlaying the four chains. 

```{R}
mcmc_trace(gpa_model_sim, size = 0.1)

```

The major criteria for a good MCMC simulation according to the trace plot is that the plot resembles white noise without overall trends. Looking at the plot above, the graphs for all the coefficients ($\beta_0,\ \beta_1,\ \beta_2,\ \sigma$) meet the criteria. 

```{r}
mcmc_dens_overlay(gpa_model_sim)

```

The results from the denisty plot agree with the trace plots. For each coefficients, even though the points for the chains do not match exactly (which is considerable given the randomness in simulation), the overall trend of the four chains are similar. 

Hence, utilizing the trace plot and density plot above, it can be assumed that the MCMC simulation for the model has adequate iterations, the markov chains are *good* and the model can be used an accurate way to approximate the posterior.  

## Interpreting the posterior (Sakaiza)

**Posterior summary statistics**

The posterior summary statistics provide us with the credible interval of each parameter.

```{r}
gpa_model_summary <- summary(gpa_model_sim)
head(as.data.frame(gpa_model_summary),-2)

```


**For Intercept**: There is a 80% posterior probability that a student that has a GPA and a SAT score of 0 exercise between 11.04 and 19.5 hours per week. 


**For SAT:** There is a 80% posterior probability that for each point increase in SAT, there is a drecrease of between 0.003 and 0	in hours of exercise per week for a student.


**For GPA:** There is a 80% posterior probability that for each point increase in GPA, there is a decrease between 1.27 and 3.30 hours in the amount of exercising a student practices per week .

These intervals are visualized by the shaded regions below: 

**Visual representation of Posterior credible interval**


```{R}
mcmc_areas(gpa_model_sim,
           pars = c("(Intercept)"),
           prob = 0.80,
           point_est="mean")

mcmc_areas(gpa_model_sim,
           pars = c("SAT"),
           prob = 0.80,
           point_est="mean")

mcmc_areas(gpa_model_sim,
           pars = c("GPA"),
           prob = 0.80,
           point_est="mean")

```


## Hypothesis testing 

Since the regression model we built provided us with 40000 alternative scenario for the trend in the relationship between Exercise, SAT and GPA, we can then compute the posterior probability that $\beta_1$ and $\beta_2$ are actually negative. 

**Data frame creation**

```{r}
gpa_model_df <- as.array(gpa_model_sim) %>%
melt %>%
pivot_wider(names_from = parameters, values_from = value)

```

**Posterior probability that $\beta_1$< 0**


```{r}
gpa_model_df %>%
  mutate(exceeds_0 = SAT < 0) %>%
  tabyl(exceeds_0)
```
Among 40 000 simulations of the same relationship, 40% display a negative correlation between SAT and exercising. 


**Posterior probability that $\beta_2$< 0**


```{r}

gpa_model_df %>%
  mutate(exceeds_0 = GPA < 0) %>%
  tabyl(exceeds_0)

```


Among 40 000 simulations, 99% display a negative correlation between GPA and exercising when controlling for the other variable. 


Because there is an 80% chance that $\beta_1$,the SAT coefficient, is between -0.003 and 0 and only 40% of the simulations demonstrate a negative correlation between these two variables, we fail to reject the null hypothesis that $\beta_1$ is equal to zero.  


Because there is an 80% chance that $\beta_2$,the GPA coefficient is between  -3.30 and  -1.27 and 99% of the simulations demonstrate a negative correlation between these two variables, we fail to reject the null hypothesis that $\beta_2$ is equal to zero. 

## Conclusion

From the analysis above we can conclude that there is not enough posterior evidence that there is a negative association or even just an association between Exercising hours of students and their SAT score, however there is ample evidence that there is a negative association between GPA and exercising hours. Using bayesian methods we went from our prior understanding of the relationship between these variables, to a posterior analysis from 40 000 simulated scenarios, which also account for uncertainties in our posterior. This was made possible by using the data a data set of 343 variables. Through this analysis, we can see the power of rtsan in simulating massive information from a small dataset.  



## References:

“Simple Normal Regression.” Bayes Rules! An Introduction to Bayesian Modeling with R, by Alicia A Johnson et al., pp. 247–264. 




